Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-05-11 paper/manuscript.Rmd
Update manuscript.Rmd
bergerud ref

Caribou were introduced to Fogo Island (n = 26 animals between 1964–67) and
Change Island (n = 5 animals in 1964) from Newfoundland as part of a series of
translocations and introductions throughout the province [@Bergerud1989].
Typical caribou predators such as black bears (*Ursus americanus*) and wolves
(*Canis lupus*) are absent from Fogo Island; the dominant predator is coyote
(*Canis latrans*), which predate calves but it remains unclear what threat they
pose, if any, to adult caribou. Given the minimal risk of predation to adults,
we propose that forage limitation is more likely to drive movement between
islands than predator avoidance. For several decades after their introduction to
Fogo Island, the caribou population did not exceed ~100 individuals (Bergerud
and Mercer 1989, Newfoundland and Labrador Wildlife Division, unpublished data).
During the 1990s, population density reached ~300 animals; anecdotal evidence
suggests caribou were first seen on nearby islands during this same period.
Increased competition for resources on Fogo due to this increase in density
could have precipitated movement to other islands.

>>>>

Caribou were introduced to Fogo Island (n = 26 animals between 1964–67) and
Change Island (n = 5 animals in 1964) from Newfoundland as part of a series of
translocations and introductions throughout the province [@Bergerud1989].
Typical caribou predators such as black bears (*Ursus americanus*) and wolves
(*Canis lupus*) are absent from Fogo Island; the dominant predator is coyote
(*Canis latrans*), which predate calves but it remains unclear what threat they
pose, if any, to adult caribou. Given the minimal risk of predation to adults,
we propose that forage limitation is more likely to drive movement between
islands than predator avoidance. For several decades after their introduction to
Fogo Island, the caribou population did not exceed ~100 individuals
[@Bergerud1989, Newfoundland and Labrador Wildlife Division, unpublished data].
During the 1990s, population density reached ~300 animals; anecdotal evidence
suggests caribou were first seen on nearby islands during this same period.
Increased competition for resources on Fogo due to this increase in density
could have precipitated movement to other islands.
